Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-10-11 main.tex
alternative algo
add def. of transitions

\newcommand{\A}{\mathcal{A}} % automata, VPA
\newcommand{\B}{\mathcal{B}} % automata (in constructions)
\newcommand{\C}{\mathcal{C}} % automata (in constructions)
\newcommand{\D}{\mathcal{D}}  % transducer = distance
\newcommand{\E}{\mathbb{E}}
%\newcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\W}{\mathbb{W}}
\newcommand{\Semiring}{\mathbb{S}}
\newcommand{\zero}{\mathbb{0}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\dom}{\ensuremath{\mathit{dom}}}
\newcommand{\seq}{\ensuremath{\mathit{seq}}}
\newcommand{\word}{\mathit{word}}

\def\SA{\textsf{sA}\xspace}
\def\WA{\textsf{wA}\xspace}
\def\SW{\textsf{sw}\xspace}
\def\swM{\textsf{swM}\xspace}
\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWTA{\textsf{swTA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\VPA{\textsf{VPA}\xspace}
\def\SVPA{\textsf{sVPA}\xspace}
\def\weight{\mathsf{weight}}
\def\mei{\mathsf{m}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phix{{\Phi_\mathsf{x}}}
\def\Phici{{\Phi_\mathsf{ci}}}
\def\Phicc{{\Phi_\mathsf{cc}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\Phicx{{\Phi_\mathsf{cx}}}
\def\wei{\mathsf{w}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\newcommand{\weie[1]}{\wei_{#1}^\mathsf{e}}
\def\weiei{\weie[\mathsf{i}]}
\def\weiec{\weie[\mathsf{c}]}
\def\weier{\weie[\mathsf{r}]}
\def\weiex{\weie[\mathsf{x}]}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
\newcommand{\rank}{\mathsf{rk}}
\newcommand{\lin}{\mathsf{lin}}
\newcommand{\tstamp}[1]{\mathpunct{:}{\scriptstyle #1}} % timestamp marker
%\newcommand{\ts}[1]{^{#1}} % timestamp marker

\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\newcommand{\ccall}[1]{\ensuremath \langle_{#1}}
\newcommand{\creturn}[1]{\ensuremath \rangle_{#1}}
%\newcommand{\creturn}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$

@@

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} \colon Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
(respectively $\mathsf{out} \colon Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
are functions defining the weight for entering
(respectively leaving) computation in a state,
and $\bar{\wei}$ is a triplet of transition functions
$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and
$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
\end{definition}
%
\noindent
We call the \emph{number of transitions} of $T$ the number of pairs of states
$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
is not the constant $\zero$.
%
\noindent
For convenience, we shall sometimes present transitions
as functions of
$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
overloading the function names,
such that, for every $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
\[
\begin{array}{rcll}
\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) &
\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
\wei_{11}(q, a, b, q') & = & \eta(a, b) &
\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
\end{array}
\]
%
\noindent
% The symbolic-weighted transducer
The \SWT $T$ computes on pairs $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$ and $t$, being respectively called \emph{input} and \emph{output} word.
\florent{It compute asynchronously: advance in $s$ and not in $t$ ($\wei_{10}$),
or the opposite ($\wei_{01}$),
or advance in both $s$ and $t$ ($\wei_{11}$).}
% \noindent
$T$ is
based on an intermediate function $\weight_T$
defined recursively, for every states $q, q' \in Q$,
and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^*$.
% \setminus \{ \< \varepsilon, \varepsilon >\}$,
%
%$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
%
\begin{align}
%\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
%\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_T(q, \varepsilon, \varepsilon, q')  & = \one
 \quad \mathrm{if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise} \label{eq:SWT-weight}\\
\weight_T(q, s, t, q') & = \displaystyle\bigoplus_{\begin{array}{c}
                                                   \scriptstyle q'' \in Q\\[-1ex]
                                                   \scriptstyle s = au,\, a \in \Sigma
                                                   \end{array}}
    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, t, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                     \scriptstyle t = bv,\, b \in \Delta\\
                                                     \end{array}}
    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', s, v, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                     \scriptstyle s = au,\, t = bv\\
                                                     \end{array}}
    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
\end{align}

@@

\noindent
 \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined  as the transducers of Definition~\ref{def:SWT},
by simply omitting the output symbols.
%
%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$,
, respectively $\mathsf{out} : Q \to \Semiring$,  %\Phi_\Sigma,$
are functions defining the weight for entering
(respectively leaving) computation in a state,
and ${\wei_1}$ is a transition function
from $Q \times Q$ into~$\Phi_{\Sigma}$.
\end{definition}
%
\noindent
%\florent{mv formal def. weight for \SWA to app.?}
As above in the case of \SWT,
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$,
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$.
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$,
The computation of $A$ on words $s \in \Sigma^*$
is based on an intermediate function $\weight_A$,
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$
%
\begin{align}
\weight_A(q, \varepsilon, q') & = \one\ \mathrm{if}\  q = q'\ 
          \mathrm{and}\  \zero\  \mathrm{otherwise}\\
%\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q}
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$
\begin{equation}
A(s)  =
\displaystyle\bigoplus_{q, q' \in Q} \init(q)
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q')
\label{eq:SWA-value}
\end{equation}

@@

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)
over  $\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, $\Semiring$ and $\bar\Phi$
is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
where $Q$ is a finite set of states,
$P$ is a finite set of stack symbols,
$\mathsf{in} : Q \to \Semiring$
(respectively $\mathsf{out} : Q \to \Semiring$)
are functions defining the weight for entering
(respectively leaving) a state,
and $\bar\wei$ is a sextuplet composed of the transition functions :
$\weii : Q \times P \times Q \to \Phici$,
$\weiei : Q \times Q \to \Phii$,
$\weic : Q \times P \times Q \times P \to \Phicc$,
$\weiec : Q \times P \times Q \to \Phic$,
$\weir : Q \times P \times Q \to \Phicr$,
$\weier : Q \times Q \to \Phir$.
%and
%$\weiex : Q \times Q \to \Phix$
%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
\end{definition}
%
As in Section~\ref{section:transducer},
we extend the above transition functions as follows
for every $q, q' \in Q$, $p \in P$,
$a \in \Deltai$,
$\call{c} \in \Deltac$,
$\return{r} \in \Deltar$,
overloading their names: % for simplicity:
\[
\begin{array}{lll}
\weii: Q \times [\Deltac \times P] \times \Deltai \times Q \to \Semiring &
\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) &
\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
%
\weiei: Q \times \Deltai \times Q \to \Semiring &
\weiei(q, a, q') = \phi_\mathsf{i}(a) &
\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
%
\weic: Q \times [\Deltac \times P] \times  [\Deltac \times P] \times Q \to \Semiring &
\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) &
\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
%
\weiec: Q \times [\Deltac \times P] \times Q \to \Semiring &
\weiec(q, {\call{c}}, p', q') = \phi_\mathsf{c}({\call{c}}) &
\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
%
\weir: Q \times [\Deltac \times P] \times \Deltar \times Q \to \Semiring &
\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) &
\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
%
\weier: Q \times \Deltar \times Q \to \Semiring &
\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
\end{array}
\]

\noindent
%The intuition is the following for the above transitions.
$\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty.
%\lydia{moved this to the beginning}
%
\noindent
$\weii$ and $\weiei$ both read an input internal symbol $a$ and change state from $q$ to $q'$,
without changing the stack.
Moreover, $\weii$ reads a pair made of
${\call{c}} \in \Deltac$ and $p \in P$ on the top of the stack
($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$).
%
\noindent
$\weic$ and $\weiec$ read the input call symbol $\call{c}'$,
push it to the stack along with $p'$, and change state from $q$ to to $q'$.
Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack
($c$ is compared to $c'$).
%and $\weiec$ applies iff the stack is empty.
%
\noindent
$\weir$ and $\weier$ read the input return symbol $\return{r}$, and change state from $q$ to to $q'$.
Moreover, $\weir$ reads and pops from stack a pair made of $\call{c}$ and $p$,
($\call{c}$ is compared to $\return{r}$).
%and $\weier$ applies iff the stack is empty.
%In this case, the weight function $\phi_\mathsf{r}$
%computes a value of matching between the call and return symbols $c$ and $r$.
%This value might be set to $\zero$ in order to express that the symbols do not match.

Formally, the transitions of the automaton~$A$ are defined
with an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
%
A configuration $q[\gamma]$
is composed of a state $q \in Q$
and a stack content $\gamma \in \Gamma^*$,
where $\Gamma = \Deltac \times P$.
Hence, $\weight_A$ is a function from
$[Q \times \Gamma^*] \times \Delta^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
The empty stack is denoted by $\bot$, and the topmost
symbol is the last pushed pair.
%
The recursive definition of $\weight_A$
enumerates each of the six possible cases:
reading $a \in \Deltai$,
%\lydia{intro to func}
or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$,
for each possible state of the stack (empty or not).
%to add to $u \in {\Delta}^*$.
%\lydia{introduced the 6 cases}
\florent{shorter notation $cp$ for $\< c, p>$ ?}

@@

>>>>

\newcommand{\A}{\mathcal{A}} % automata, VPA
\newcommand{\B}{\mathcal{B}} % automata (in constructions)
\newcommand{\C}{\mathcal{C}} % automata (in constructions)
\newcommand{\D}{\mathcal{D}}  % transducer = distance
\newcommand{\E}{\mathbb{E}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\G}{\mathcal{G}}
%\newcommand{\P}{\mathcal{P}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\W}{\mathbb{W}}
\newcommand{\Semiring}{\mathbb{S}}
\newcommand{\zero}{\mathbb{0}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\dom}{\ensuremath{\mathit{dom}}}
\newcommand{\seq}{\ensuremath{\mathit{seq}}}
\newcommand{\word}{\mathit{word}}

\def\SA{\textsf{sA}\xspace}
\def\WA{\textsf{wA}\xspace}
\def\SW{\textsf{sw}\xspace}
\def\swM{\textsf{swM}\xspace}
\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWTA{\textsf{swTA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\VPA{\textsf{VPA}\xspace}
\def\SVPA{\textsf{sVPA}\xspace}
\def\weight{\mathsf{weight}}
\def\mei{\mathsf{m}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phix{{\Phi_\mathsf{x}}}
\def\Phici{{\Phi_\mathsf{ci}}}
\def\Phicc{{\Phi_\mathsf{cc}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\Phicx{{\Phi_\mathsf{cx}}}
\def\wei{\mathsf{w}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\newcommand{\weie[1]}{\wei_{#1}^\mathsf{e}}
\def\weiei{\weie[\mathsf{i}]}
\def\weiec{\weie[\mathsf{c}]}
\def\weier{\weie[\mathsf{r}]}
\def\weiex{\weie[\mathsf{x}]}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
\newcommand{\rank}{\mathsf{rk}}
\newcommand{\lin}{\mathsf{lin}}
\newcommand{\tstamp}[1]{\mathpunct{:}{\scriptstyle #1}} % timestamp marker
%\newcommand{\ts}[1]{^{#1}} % timestamp marker
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\newcommand{\ccall}[1]{\ensuremath \langle_{#1}}
\newcommand{\creturn}[1]{\ensuremath \rangle_{#1}}
%\newcommand{\creturn}[1]{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\newcommand{\src}{\mathsf{src}}
\newcommand{\trg}{\mathsf{trg}}

@@

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} \colon Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
(respectively $\mathsf{out} \colon Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
are functions defining the weight for entering
(respectively leaving) computation in a state,
and $\bar{\wei}$ is a triplet of transition functions
$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and
$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
\end{definition}
%
\noindent
%We call the \emph{number of transitions} of $T$ the number of pairs of states
%$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
%is not the constant $\zero$.
%
\noindent
A tuple of $Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$
is called a \emph{transition} of $T$.
For convenience, we shall sometimes present the above $\wei_{10}$, $\wei_{01}$, $\wei_{11}$
%(overloading the function names),
as mappings associating to every transition a weight value in $\Semiring$ as follows,
for every $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
\[
\begin{array}{rcll}
\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) &
\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
\wei_{11}(q, a, b, q') & = & \eta(a, b) &
\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
\end{array}
\]
%
\noindent
% The symbolic-weighted transducer
The \SWT $T$ computes on pairs $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$ and $t$, being respectively called \emph{input} and \emph{output} word.
\florent{It compute asynchronously: advance in $s$ and not in $t$ ($\wei_{10}$),
or the opposite ($\wei_{01}$),
or advance in both $s$ and $t$ ($\wei_{11}$).}
% \noindent
$T$ is
based on an intermediate function $\weight_T$
defined recursively, for every states $q, q' \in Q$,
and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^*$.
% \setminus \{ \< \varepsilon, \varepsilon >\}$,
%
%$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
%
\begin{align}
%\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
%\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_T(q, \varepsilon, \varepsilon, q')  & = \one
 \quad \mathrm{if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise} \label{eq:SWT-weight}\\
\weight_T(q, s, t, q') & = \displaystyle\bigoplus_{\begin{array}{c}
                                                   \scriptstyle q'' \in Q\\[-1ex]
                                                   \scriptstyle s = au,\, a \in \Sigma
                                                   \end{array}}
    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, t, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                     \scriptstyle t = bv,\, b \in \Delta\\
                                                     \end{array}}
    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', s, v, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                     \scriptstyle s = au,\, t = bv\\
                                                     \end{array}}
    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
\end{align}

@@

\noindent
 \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined  as the transducers of Definition~\ref{def:SWT},
by simply omitting the output symbols.
%
%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$,
, respectively $\mathsf{out} : Q \to \Semiring$,  %\Phi_\Sigma,$
are functions defining the weight for entering
(respectively leaving) computation in a state,
and ${\wei_1}$ is a transition function
from $Q \times Q$ into~$\Phi_{\Sigma}$.
\end{definition}
%
\noindent
%\florent{mv formal def. weight for \SWA to app.?}
A \SWA \emph{transition} is a triplet of $Q \times \Sigma \times Q$, 
and, as in the case of \SWT,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$.
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$,
The computation of $A$ on words $s \in \Sigma^*$
is based on an intermediate function $\weight_A$,
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$
%
\begin{align}
\weight_A(q, \varepsilon, q') & = \one\ \mathrm{if}\  q = q'\ 
          \mathrm{and}\  \zero\  \mathrm{otherwise}\\
%\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q}
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$
\begin{equation}
A(s)  =
\displaystyle\bigoplus_{q, q' \in Q} \init(q)
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q')
\label{eq:SWA-value}
\end{equation}

@@

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA)
over  $\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, $\Semiring$ and $\bar\Phi$
is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
where $Q$ is a finite set of states,
$P$ is a finite set of stack symbols,
$\mathsf{in} : Q \to \Semiring$
(respectively $\mathsf{out} : Q \to \Semiring$)
are functions defining the weight for entering
(respectively leaving) a state,
and $\bar\wei$ is a sextuplet composed of the transition functions :
$\weii : Q \times P \times Q \to \Phici$,
$\weiei : Q \times Q \to \Phii$,
$\weic : Q \times P \times Q \times P \to \Phicc$,
$\weiec : Q \times P \times Q \to \Phic$,
$\weir : Q \times P \times Q \to \Phicr$,
$\weier : Q \times Q \to \Phir$.
%and
%$\weiex : Q \times Q \to \Phix$
%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
\end{definition}
%
As in Section~\ref{section:transducer},
we extend the above transition functions as follows
for every $q, q' \in Q$, $p \in P$,
$a \in \Deltai$,
$\call{c} \in \Deltac$,
$\return{r} \in \Deltar$,
overloading their names: % for simplicity:
\[
\begin{array}{lll}
\weii: Q \times [\Deltac \times P] \times \Deltai \times Q \to \Semiring &
\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) &
\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
%
\weiei: Q \times \Deltai \times Q \to \Semiring &
\weiei(q, a, q') = \phi_\mathsf{i}(a) &
\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
%
\weic: Q \times [\Deltac \times P] \times  [\Deltac \times P] \times Q \to \Semiring &
\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) &
\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
%
\weiec: Q \times [\Deltac \times P] \times Q \to \Semiring &
\weiec(q, {\call{c}}, p', q') = \phi_\mathsf{c}({\call{c}}) &
\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
%
\weir: Q \times [\Deltac \times P] \times \Deltar \times Q \to \Semiring &
\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) &
\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
%
\weier: Q \times \Deltar \times Q \to \Semiring &
\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
\end{array}
\]
The tuples in the definition domains of the above functions 
are also called \emph{transitions} of the \SWVPA.
%
\noindent
%The intuition is the following for the above transitions.
Intuitively, $\weiei$, $\weiec$, and $\weier$ describe the cases where the stack is empty.
The transitions of $\weii$ and $\weiei$ both read an internal symbol~$a$ 
and change state from $q$ to $q'$, without changing the stack.
Moreover, $\weii$ reads a pair made of
${\call{c}} \in \Deltac$ and $p \in P$ on the top of the stack
($c$ is compared to $a$ by the weight function $\eta_\mathsf{ci} \in \Phici$).
%
\noindent
The transitions of $\weic$ and $\weiec$ read the call symbol $\call{c}'$,
push it to the stack along with $p'$, and change state from $q$ to to $q'$.
Moreover, $\weic$ reads ${\call{c}}$ and $p$ at the top of the stack
($c$ is compared to $c'$).
%and $\weiec$ applies iff the stack is empty.
%
\noindent
Finally, $\weir$ and $\weier$ read the return symbol $\return{r}$, and change state from $q$ to to $q'$.
Moreover, $\weir$ reads and pops from stack a pair made of $\call{c}$ and $p$
($\call{c}$ is compared to $\return{r}$).
%and $\weier$ applies iff the stack is empty.
%In this case, the weight function $\phi_\mathsf{r}$
%computes a value of matching between the call and return symbols $c$ and $r$.
%This value might be set to $\zero$ in order to express that the symbols do not match.

Formally, the computations of the automaton~$A$ are defined
with an intermediate function $\weight_A$, like in Section~\ref{sec:SWT}.
%
A configuration $q[\gamma]$
is composed of a state $q \in Q$
and a stack content $\gamma \in \Gamma^*$,
where $\Gamma = \Deltac \times P$.
Hence, $\weight_A$ is a function from
$[Q \times \Gamma^*] \times \Delta^* \times [Q \times \Gamma^*]$ into~$\Semiring$.
The empty stack is denoted by $\bot$, and the topmost
symbol is the last pushed pair.
%
The recursive definition of $\weight_A$
enumerates each of the six possible cases:
reading $a \in \Deltai$,
%\lydia{intro to func}
or $\call{c} \in \Deltac$, or $\return{r} \in \Deltar$,
for each possible state of the stack (empty or not).
%to add to $u \in {\Delta}^*$.
%\lydia{introduced the 6 cases}
\florent{shorter notation $cp$ for $\< c, p>$ ?}

@@

\input{altalgo}
